---
title: "MA22004 Course Design Plan"
author: "Dr Eric Hall [(ehall001@dundee.ac.uk)](mailto:ehall001@dundee.ac.uk)"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  bookdown::pdf_book:
    keep_tex: yes
  bookdown::gitbook:
    config:
      toc:
        collapse: no
        scroll_highlight: yes
      toolbar:
        position: fixed
      download: pdf
      sharing: null
      fontsettings:
        theme: white
        family: sans
        size: 2
      info: yes
      highlight: pygments
      number_sections: no
description: Notes for course design, MA22004 Statistics II, University of Dundee.
---

```{r setup, include=FALSE}
#  bookdown::render_book("index.Rmd", "bookdown::gitbook", output_dir = "docs/")

# packages
library(bookdown)
library(plotly)
library(tidyverse)
library(reshape2)
library(latex2exp)
library(svglite)
library(knitr)

knitr::opts_chunk$set(echo = TRUE, comment = "", dev = "svglite") #, dev = "svglite"
#plotly::config(plot_ly(), displaylogo = FALSE, mathjax = "cdn")
```



# Learning {#learning}

> *Wir shaffen das.*

*Summary:* runs for 11 weeks (20 SCQF/10 ECTS) approx 200 hours student effort including 55 contact hours.

There will be 1 seminars of one-hour each (**timetabled, online**), 1 workshop of one-hour (**timetabled, face-to-face**), and 1 lab of two hours (**online, asynchronous**). The expectation will be that students access and engage with the lecture notes and curated digital content before participating in the seminar. 

## Asynchronous content {-}

All of this content will be available online. 

Course notes (ACQUIRING)
:  **needs to be prepared** (use of existing hand-written notes with some modificaitons), this content will consist of HTML and PDF notes generated using `Rmarkdown`.

Interactive computer labs (PRODUCING, PRACTICING)
:  **needs to be prepared** (use of prepared labs with content converted to `Rmarkdown` and interactive elements added), this content will consist of  interactive `R` computer labs coded using `learnr` and `gradethis`. **Digital [need](#needs):** RStudio Cloud.

Curated digital content (INQUIRING)
:  a collection of digital content for students to explore, e.g., online lectures by [David Spiegelhalter](https://www.youtube.com/watch?v=a1PtQ67urG4), [Hans Rosling](https://www.youtube.com/watch?v=hVimVzgtD6w), etc; [spurious correlations](http://tylervigen.com/spurious-correlations), [Datasaurus Dozen]([https://www.autodeskresearch.com/publications/samestats), etc. Shared via Blackboard or [padlet](https://padlet.com/).

## Synchronous content {-}

This content must be **timetabled**.

### Face-to-face {-}

Workshops (PRACTICING, COLLABORATING, DISCUSSING)
:  In which students will meet together to work in small groups (5 or 6). The first couple of meetings will involve meeting outdoors, e.g., UoD Botanical Gardens or a park, to collect data (circumference of trees, heights of people, number of people crossing at a busy intersection, etc) that will later be discussed an analyzed. The students will utilized worksheet handouts that are already prepared.

### Online {-}

Seminars (DISCUSSING, PRODUCING, PRACTICING)
:  In which we will have a discussion of the main topics for the week, guided by student polling (using either [mentimeter](https://www.mentimeter.com/) or [itempool](https://itempool.com/). Periodic quizzes will be given in seminars to assess student attainment. A portion of each seminar will involve a lab presentation given by one of the groups.

Examinations (PRODUCING)
:  Midterm and Final examinations will need to be timetabled to lessen the chance of cheating. The Final Examination can also be given during the May diet.

# Engagement {#engagement}

> The active learning categorization for each teaching task are identified in [ยง Learning](#learning)

There are several points that can be measured to assess engagement, including: 

+ Students will be presented with a new collection of notes that will roll out weekly using a tool like [perusall](https://perusall.com/) (or possibly Blackboard forums). They will be responsible for posting both questions and responses to the message boards for each new collection of notes.  
+ During seminars, students will be presented with **polls** to encourage reflection and discussion.
+ During seminars, students will take turns giving a lab report **presentation** with their workshop groups. This will encourage students to challenge each others analysis.
+ Students will be expected to hand-in assignments for continuous assessment.
+ Students will be expected to sit Midterm examinations. The grades of these examinations will offer opportunities for intervention. 

# Attainment



# Assessment {#assessment}

The module assessment will be weighted:
```{r table, echo=FALSE}
df <- tibble(Assessment=c("Assignments", "Midterm Exam 1", "Midterm Exam 2", "Final Exam"),
             Weight=c("20%", "20%", "20%", "40%"))
kable(df)
```
Note: This is a modification from the previous split (60% degree exam, 20% two 
class tests, and 20% coursework). As the students will not be given a revision period, the Final Exam will need to be less cumulative in scope and therefore additional weight will be placed on the Midterm Exams.

## Examinations {-}
The **Final Exam** will be a two-hour hand-written exam. Students will be given one additional hour to scan, code, and upload their manuscripts using [gradescope](https://www.gradescope.com/). This process will be thoroughly discussed and practiced by the students with a dummy exam in advance of a real submission. The Final Exam will either be given in the last week of term (week 11) or during the May exam diet.

The **Midterm Exams** will either be hand-written or computer-assessed and will be one-hour in scope instead of two. These will likely be held in week 4 and week 8. 

## Assignments {-}

The assessed courwork will include 6 lab reports and 1 lab report presentation (group work). 

# NEEDS {#needs}

## Digital needs {-}
1. bringing computer lab into cloud: semester subsription to [RStudio Cloud](https://rstudio.cloud/)^[A better long-term solution would be to run `RStudio Server Pro` and `RStudio Connect` on in-house server running, licenses are available for [free](https://rstudio.com/pricing/academic-pricing/) for teaching purposes.] (approx. 1000 USD)
2. polling: [mentimeter](https://www.mentimeter.com/) or [itempool](https://itempool.com/)

## Hardware needs {-}
1. Better microphone
2. Tablet with stylus


# Rmarkdown Example {-}

## Math {-}

Here is an example page containing math markdown.

```{latex} 
We can use inline math, $f_X$, $f_Y$, and $f_{X,Y}$, as well as display math, 
\[
  I(X;Y)
  = \iint_{\mathcal{X} \otimes \mathcal{Y}}
  \log\left( \frac{f_{X,Y}(x,y)}{f_{X}(x)f_{Y}(y)} \right)
  f_{X,Y}(x,y) \mathrm{d}y \mathrm{d}x\,.
\]
```

to generate:

We can use inline math, $f_X$, $f_Y$, and $f_{X,Y}$, as well as display math, 
\[
  I(X;Y)
  = \iint_{\mathcal{X} \otimes \mathcal{Y}}
  \log\left( \frac{f_{X,Y}(x,y)}{f_{X}(x)f_{Y}(y)} \right)
  f_{X,Y}(x,y) \mathrm{d}y \mathrm{d}x\,.
\]

## Plots {-}

```{r normals-same-var, echo=TRUE, fig.cap="Two normal variates with different means and same variance. Note this figure is a scalable vector graphic --- from what I understand, this is better from an accessbility standpoint."}
x <- seq(-8, 8, length=1000)
y0 <- dnorm(x, -2, 1)
y1 <- dnorm(x, 2, 1)
df <- tibble(x, y0, y1)
df <- melt(df, id.var = "x", value.name = "y")

ggplot(data = df, aes(x = x, color = variable)) + geom_line(aes(y=y)) 
```

The figure with caption caption is created by typing the code directly into the markdown document:
````
```{r normals-same-var, echo=TRUE, fig.cap="Two normal variates with different means and same variance. Note this figure is a scalable vector graphic --- from what I understand, this is better from an accessbility standpoint."}`r ''`
  x <- seq(-8, 8, length=1000)
  y0 <- dnorm(x, -2, 1)
  y1 <- dnorm(x, 2, 1)
  df <- tibble(x, y0, y1)
  df <- melt(df, id.var = "x", value.name = "y")
  ggplot(data = df, aes(x = x, color = variable)) + geom_line(aes(y=y)) 
```
````